---
title: "Stochastic Method in Finance"
subtitle: "Assignement: Exotic Options"
author: "Antoine Marie - 24-613-762"
format: 
  html: 
    embed-resources: true  
    self-contained: true
    self-contained-math: true
    math-method: mathjax
    
highlight-style: dracula
editor: 
  markdown: 
    wrap: 72
---
```{r setup, echo = FALSE, eval = TRUE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, eval = TRUE)
```

# R Setup
```{r}
# loading the library
library(readxl)
library(dplyr)
library(ggplot2)
```

# 0) Data Preparation

##── 1. Download Apple Data 

```{r}
# reading the file
getwd()
data <- read_excel("AAPL.xlsx")
```
Added to clean the environment:
rm(list = ls())

# ── 2. Cleaning 

```{r}
# Convert date to Date format
data$Date <- as.Date(data$Date)

# Rename AAPL.US to Price
colnames(data)[colnames(data) == "AAPL.US"] <- "Price"

# Convert Price to numeric (in case it was read as character)
data$Price <- as.numeric(data$Price)

# Sort ascending (oldest first)
data <- data[order(data$Date), ]

# Filter to 2020 onwards
data <- data[data$Date >= as.Date("2020-01-01"), ]

# Drop unnecessary columns (keep only Date and Price)
data <- data[, c("Date", "Price")]

# Sum the NA 
sum(is.na(data))

# Remove any rows with NA (if they were any)
data <- na.omit(data)
# in our case there is none so it doesn't remove anything
```
# ── 3. Verification 
```{r}
cat("Date range:  ", format(min(data$Date)), "to", format(max(data$Date)), "\n")
cat("Observations:", nrow(data), "\n")
cat("Date class:  ", class(data$Date), "\n")
cat("Price class: ", class(data$Price), "\n")
cat("Any NAs:     ", anyNA(data), "\n")
# small excerpt of the data
head(data)
tail(data)
```

---

## i) Binomial Tree

# ── Given Parameters 
```{r}
n <- 25           # number of periods
r <- 0.01         # interest rate per annum
S0 <- 267.6100    # Close value on the 27th April 2026 
p <- 0.5          # probability of a up or down
T <- 6/12         # anualise the maturity
delta_t <- T/n    # go from discrete to continuous Random walk to Brownian motion

```

# ── Computations of parameters 
```{r}
#Daily log returns 
returns <- diff(log(data$Price)) #Calculate the daily log returns, slide 

sigma_daily <- sd(returns, na.rm = TRUE) #Calculate the standard deviation, remove na values (the one where we could not compute the log return (the first observation has no daily log return))

# sort the data by year and extract the number of trading day for each year
trading_days_per_year <- table(format(data$Date, "%Y"))
# do the mean for all the year
avg_trading_days <- mean(trading_days_per_year)
# compute the sigma which should be lower then the initial sigma computed on 250 because be do not use all the day in 2026 ( for example: 250,250,250,250,120)
sigma_lower <- sigma_daily * sqrt(avg_trading_days)

#calculate the annualized volatility, given in assignment (take 250 days as a given)
sigma <- sigma_daily * sqrt(250) 

#Up and down factors 
u <- exp(sigma * sqrt(delta_t))  #page 2
d <- exp(-sigma * sqrt(delta_t))
```

# ── The binomial tree (matrix format) 
```{r}
# Initialize price tree (# creation of the matrix)
stock_tree <- matrix(0, nrow = n + 1, ncol = n + 1)

for (i in 0:n) {        # i = time step
  for (j in 0:i) {      # j = number of up moves
    stock_tree[i - j + 1, i + 1] <- S0 * u^j * d^(i - j)
  }
}

print(stock_tree)
#middle node at maturity [13;25] = S0

# All non-zero terminal nodes
terminal <- stock_tree[stock_tree[, n + 1] != 0, n + 1]

cat("# of terminal nodes:", length(terminal), "\n")
cat("Maximum price:           ", max(terminal), "\n")
cat("Minimum price:           ", min(terminal), "\n")
cat("Middle node (~S0):       ", median(terminal), "\n")

# All terminal prices
cat("\nAll terminal prices:\n")
print(round(terminal, 2))

# Binonmial tree visual 
tree_nodes <- data.frame()
for (i in 0:4) {
  for (j in 0:i) {
    price       <- S0 * u^j * d^(i - j)
    is_reconcile <- (i == 2 & j == 1)
    tree_nodes  <- rbind(tree_nodes, data.frame(
      t         = i,
      y         = j - i/2,
      price     = round(price, 2),
      reconcile = is_reconcile
    ))
  }
}

tree_edges <- data.frame()
for (i in 0:3) {
  for (j in 0:i) {
    y_current <- j - i/2
    tree_edges <- rbind(tree_edges, data.frame(
      x = i, xend = i + 1,
      y = y_current, yend = (j + 1) - (i + 1)/2
    ))
    tree_edges <- rbind(tree_edges, data.frame(
      x = i, xend = i + 1,
      y = y_current, yend = j - (i + 1)/2
    ))
  }
}

# draw the tree
p_tree <- ggplot() +
  geom_segment(data = tree_edges,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "grey40", linewidth = 0.6) +
  geom_point(data = tree_nodes[!tree_nodes$reconcile, ],
             aes(x = t, y = y),
             size = 16, color = "grey40") +
  geom_point(data = tree_nodes[tree_nodes$reconcile, ],
             aes(x = t, y = y),
             size = 16, color = "red") +
  geom_text(data = tree_nodes,
            aes(x = t, y = y, label = paste0("$", price)),
            color = "white", size = 2.8, fontface = "bold") +
  annotate("text",
           x = tree_nodes$t[tree_nodes$reconcile] + 0.12,
           y = tree_nodes$y[tree_nodes$reconcile] - 0.42,
           label = "S0 * u * d = S0",
           color = "red", size = 3.2, hjust = 0) +
  annotate("text",
           x = 0:4, y = rep(-2.9, 5),
           label = paste0("t = ", 0:4),
           size = 3.2, color = "grey40") +
  scale_x_continuous(limits = c(-0.3, 5.0)) +
  scale_y_continuous(limits = c(-3.2, 2.6)) +
  labs(
    title    = "Binomial Tree — First 4 Periods",
    subtitle = paste0("S0 = $", round(S0, 2),
                      "   u = ", round(u, 4),
                      "   d = ", round(d, 4),
                      "   p = ", p),
    caption  = "Red node: recombination S0 * u * d = S0"
  ) +
  theme_void() +
  theme(
    plot.title    = element_text(size = 14, face = "bold",
                                 hjust = 0.5, margin = margin(b = 5)),
    plot.subtitle = element_text(size = 10, hjust = 0.5,
                                 color = "grey40", margin = margin(b = 15)),
    plot.caption  = element_text(size = 9, hjust = 0.5,
                                 color = "red", margin = margin(t = 10)),
    plot.margin   = margin(15, 15, 15, 15)
  )

print(p_tree)
```

---

## ii) Compute the average stock price for each terminal node of the tree 

```{r}
# Using forward induction to have cumulative sum C_t along each path, chapter 2 slide 28
#2^25 node to compute = way too much for R -> 

states <- data.frame(
  S      = S0,      # current stock price
  C      = S0,      # cumulative sum of price (only S0 so far)
  weight = 1.0      # physical probability weight (p = 0.5)
)

for (step in 1:n) {
  up <- data.frame(
    S      = states$S * u,
    C      = states$C + states$S * u,
    weight = states$weight * p
  )
  down <- data.frame(
    S      = states$S * d,
    C      = states$C + states$S * d,
    weight = states$weight * (1 - p)
  )
  states <- rbind(up, down)
}

# Compute arithmetic average for each terminal path
states$S_bar <- states$C / (n + 1)

cat("\n=== Task (ii): Arithmetic Average Summary ===\n")
cat("Number of terminal paths:", nrow(states), "\n")
cat("Max average price:  $", round(max(states$S_bar), 4), "\n")
cat("Min average price:  $", round(min(states$S_bar), 4), "\n")
cat("Mean of averages:   $", round(mean(states$S_bar), 4), "\n")

averages_list <- sort(states$S_bar, decreasing = TRUE)


# Plot 2: Distribution of Terminal prices, Using binomial formula with p = 0.5 (physical probability)

j_vals   <- 0:n
prices_T <- S0 * u^j_vals * d^(n - j_vals)
probs_p  <- choose(n, j_vals) * (0.5)^n

terminal_df <- data.frame(
  price = prices_T,
  prob  = probs_p
)

p_dist <- ggplot(terminal_df, aes(x = price, y = prob)) +
  geom_col(fill = "grey40", alpha = 0.85, width = diff(range(prices_T))/30) +
  geom_vline(xintercept = S0,
             color = "red", linetype = "dashed", linewidth = 0.8) +
  annotate("text",
           x = S0 + diff(range(prices_T)) * 0.03,
           y = max(probs_p) * 0.93,
           label = paste0("S0 = $", round(S0, 2)),
           color = "red", size = 3.5, hjust = 0) +
  labs(
    title    = "Distribution of Terminal Stock Prices at t = n",
    subtitle = paste0("n = ", n, " periods   |  Physical Probability p = 0.5"),
    x        = "Terminal Stock Price ($)",
    y        = "Probability"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 10, hjust = 0.5, color = "grey40"),
    axis.title       = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

print(p_dist)

# PLOT 3: DISTRIBUTION OF TERMINAL AVERAGES ─────────────

p_avg <- ggplot(states, aes(x = S_bar, weight = weight)) +
  geom_histogram(bins = 50, fill = "grey40", alpha = 0.85, color = "white") +
  geom_vline(xintercept = S0,
             color = "red", linetype = "dashed", linewidth = 0.8) +
  annotate("text",
           x = S0 + 1, y = Inf, vjust = 2,
           label = paste0("S\u2080 = $", round(S0, 2)),
           color = "red", size = 3.5, hjust = 0) +
  labs(
    title    = "Distribution of Arithmetic Averages at Maturity",
    subtitle = paste0(nrow(states), " paths   |   p = 0.5"),
    x        = "Arithmetic Average Price ($)",
    y        = "Number of Paths"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 10, hjust = 0.5, color = "grey40"),
    axis.title       = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

print(p_avg)
```

The forward induction works by recognizing that the cumulative sum $C_t = \sum_{j=0}^{t} S_j$ satisfies the simple recursion $C_t = C_{t-1} + S_t$, which means we only need to track the pair $(S_t, C_t)$ at each step rather than the full path history, it is the Markov property of the augmented state space. At each time step, every existing state branches into exactly two children (up and down), carrying forward both the new stock price and the updated cumulative sum, with probability weights that multiply by $p = 0.5$ at each branch. After $n = 25$ steps we have exactly $2^{25}$ states, each representing one unique path through the tree, and the arithmetic average is immediately recovered as $\bar{S}_n = \frac{C_n}{(n+1)}$,  no path information is lost.

---

## iii) Compute the different payoffs.

Define the risk-neutral probability and the discount factor using discrete compounding (because of the binomial model)

```{r set-up function parameter}
# Need to compute the risk neutral probability under which there is no arbitrage opportunity (Fundamental Theorem of Asset Pricing)

# define the one-step risk-free growth factor
R_period <- 1 + (r*delta_t)
# define the risk-neutral probability
q <- (R_period-d)/(u - d)
# define the discount factor 
discount <- 1 / R_period

# print recap of all value needed for the rest of the code
cat("Daily volatility:         ", sigma_daily, "\n")
cat("Annual volatility:        ", sigma, "\n")
cat("Up factor u:              ", u, "\n")
cat("Down factor d:            ", d, "\n")
cat("One-step growth R_period: ", R_period, "\n")
cat("Risk-neutral prob. q:     ", q, "\n")
cat("Discount factor:          ", discount, "\n")

```

```{r compute the payoff}
# we need to compute the payoff at maturity for every path.
states$payoff <- pmax(states$S - states$S_bar, 0)

# print the recap of the payoff ___________________________
cat("\n=== Task (iii): Floating-Strike Asian Call Payoff ===\n")
cat("Number of terminal paths: ", nrow(states), "\n")
cat("Minimum payoff:          $", round(min(states$payoff), 4), "\n")
cat("Maximum payoff:          $", round(max(states$payoff), 4), "\n")
cat("Mean payoff:             $", round(mean(states$payoff), 4), "\n")
cat("Share positive payoffs:   ", round(mean(states$payoff > 0) * 100, 2), "%\n")

# Plot the distribution of the payoff at maturity _________
ggplot(states, aes(x = payoff)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white", alpha = 0.8) +
  labs(
    title = "Distribution of Potential Payoffs at Maturity",
    subtitle = "Unweighted across all terminal paths",
    x = "Payoff at maturity",
    y = "Frequency"
  ) +
  theme_minimal()



head(states)
```

## i.v) Weighted Outcome under risk-neutral probability
Use the weighted probability under the no-arbitrage opportunity assumption. 


## v) Estimate the pricing option: (use backward induction)


```{r option value using backward induction}
# we have each terminal payoff
values <- states$payoff 

# in order to use backward induction we need to start from maturity and then loops through each eateriation until we reach time 0 (starting point)
for (step in n:1) {
  half   <- length(values) / 2
  v_up   <- values[1:half]
  v_down <- values[(half + 1):length(values)]
  values <- discount * (q * v_up + (1 - q) * v_down)
}

V0_b<- values
cat(" Option price at t = 0 with backward induction: $", round(V0_b,4))

```
```{r option value using direct forward calculation}
# find the number of up moves for each path
states$up_moves <- round((log(states$S / S0) - n * log(d)) / (log(u) - log(d))) 

# Risk-neutral probability of each individual path
states$rn_weight <- q^states$up_moves * (1 - q)^(n - states$up_moves)

# Expected payoff under the risk-neutral measure
expected_payoff <- sum(states$rn_weight * states$payoff)

# Discount the expected payoff back to time 0
V0_f<- discount^n * expected_payoff

cat(" The option price at T = 0 with direct forward calculation: $", round(V0_f,4))

```


```{r comparing both the backward induction with direct forward computation}
cat("Difference between methods used:\n")
cat("V0 (backward induction):", round(V0_b, 10), "\n")
cat("V0 (direct calculation):", round(V0_f, 10), "\n")
```



