---
title: "Stochastic Method in Finance"
subtitle: "Assignement: Exotic Options"
author: "Antoine Marie - 24-613-762"
format: 
  html: 
    embed-resources: true  
    self-contained: true
    self-contained-math: true
    math-method: mathjax
    
highlight-style: dracula
editor: 
  markdown: 
    wrap: 72
---
```{r setup, echo = FALSE, eval = TRUE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, eval = TRUE)
```

# R Setup
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
```

# 0) Data Preparation

##── 1. Download Apple Data 

```{r}
getwd()
setwd("C:/William/University_GitHub/Stochastic Method in Finance/Assignment")
data <- read_excel("AAPL.xlsx")
```
Added to clean the environment:
rm(list = ls())

# ── 2. Cleaning 

```{r}
# Convert date to Date format
data$Date <- as.Date(data$Date)

# Rename AAPL.US to Price
colnames(data)[colnames(data) == "AAPL.US"] <- "Price"

# Convert Price to numeric (in case it was read as character)
data$Price <- as.numeric(data$Price)

# Sort ascending (oldest first)
data <- data[order(data$Date), ]

# Filter to 2020 onwards
data <- data[data$Date >= as.Date("2020-01-01"), ]

# Drop unnecessary columns (keep only Date and Price)
data <- data[, c("Date", "Price")]

# Remove any rows with NA
data <- na.omit(data)
```
# ── 3. Verification 
```{r}
cat("Date range:  ", format(min(data$Date)), "to", format(max(data$Date)), "\n")
cat("Observations:", nrow(data), "\n")
cat("Date class:  ", class(data$Date), "\n")
cat("Price class: ", class(data$Price), "\n")
cat("Any NAs:     ", anyNA(data), "\n")

head(data)
tail(data)
```

---

## i) Binomial Tree

# ── Given Parameters 
```{r}
n <- 25           # number of periods
r <- 0.01         # interest rate per annum
S0 <- 267.6100    # Close value on the 27th April 2026 
p <- 0.5          # probability of a up or down
T <- 6/12         # anualise the maturity
delta_t <- T/n    # go from discrete to continuous Random walk to Brownian motion

```

# ── Computations of parameters 
```{r}
#Daily log returns 
returns <- diff(log(data$Price)) #Calculate the daily log returns, slide 

sigma_daily <- sd(returns, na.rm = TRUE) #Calculate the standard deviation, remove na values

sigma <- sigma_daily * sqrt(250) #calculate the annualized volatility, given in assignment

#Up and down factors 
u <- exp(sigma * sqrt(delta_t))     #page 2
d <- exp(-sigma * sqrt(delta_t))

```

# ── The binomial tree (matrix format) 
```{r}
# Initialize price tree
stock_tree <- matrix(0, nrow = n + 1, ncol = n + 1)

for (i in 0:n) {        # i = time step
  for (j in 0:i) {      # j = number of up moves
    stock_tree[i - j + 1, i + 1] <- S0 * u^j * d^(i - j)
  }
}

print(stock_tree)
#middle node at maturity [13;25] = S0

# All non-zero terminal nodes
terminal <- stock_tree[stock_tree[, n + 1] != 0, n + 1]

cat("# of terminal nodes:", length(terminal), "\n")
cat("Maximum price:           ", max(terminal), "\n")
cat("Minimum price:           ", min(terminal), "\n")
cat("Middle node (~S0):       ", median(terminal), "\n")

# All terminal prices
cat("\nAll terminal prices:\n")
print(round(terminal, 2))

# Binonmial tree visual 
tree_nodes <- data.frame()
for (i in 0:4) {
  for (j in 0:i) {
    price       <- S0 * u^j * d^(i - j)
    is_reconcile <- (i == 2 & j == 1)
    tree_nodes  <- rbind(tree_nodes, data.frame(
      t         = i,
      y         = j - i/2,
      price     = round(price, 2),
      reconcile = is_reconcile
    ))
  }
}

tree_edges <- data.frame()
for (i in 0:3) {
  for (j in 0:i) {
    y_current <- j - i/2
    tree_edges <- rbind(tree_edges, data.frame(
      x = i, xend = i + 1,
      y = y_current, yend = (j + 1) - (i + 1)/2
    ))
    tree_edges <- rbind(tree_edges, data.frame(
      x = i, xend = i + 1,
      y = y_current, yend = j - (i + 1)/2
    ))
  }
}

p_tree <- ggplot() +
  geom_segment(data = tree_edges,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "grey40", linewidth = 0.6) +
  geom_point(data = tree_nodes[!tree_nodes$reconcile, ],
             aes(x = t, y = y),
             size = 16, color = "grey40") +
  geom_point(data = tree_nodes[tree_nodes$reconcile, ],
             aes(x = t, y = y),
             size = 16, color = "red") +
  geom_text(data = tree_nodes,
            aes(x = t, y = y, label = paste0("$", price)),
            color = "white", size = 2.8, fontface = "bold") +
  annotate("text",
           x = tree_nodes$t[tree_nodes$reconcile] + 0.12,
           y = tree_nodes$y[tree_nodes$reconcile] - 0.42,
           label = "S0 * u * d = S0",
           color = "red", size = 3.2, hjust = 0) +
  annotate("text",
           x = 0:4, y = rep(-2.9, 5),
           label = paste0("t = ", 0:4),
           size = 3.2, color = "grey40") +
  scale_x_continuous(limits = c(-0.3, 5.0)) +
  scale_y_continuous(limits = c(-3.2, 2.6)) +
  labs(
    title    = "Binomial Tree — First 4 Periods",
    subtitle = paste0("S0 = $", round(S0, 2),
                      "   u = ", round(u, 4),
                      "   d = ", round(d, 4),
                      "   p = ", p),
    caption  = "Red node: recombination S0 * u * d = S0"
  ) +
  theme_void() +
  theme(
    plot.title    = element_text(size = 14, face = "bold",
                                 hjust = 0.5, margin = margin(b = 5)),
    plot.subtitle = element_text(size = 10, hjust = 0.5,
                                 color = "grey40", margin = margin(b = 15)),
    plot.caption  = element_text(size = 9, hjust = 0.5,
                                 color = "red", margin = margin(t = 10)),
    plot.margin   = margin(15, 15, 15, 15)
  )

print(p_tree)
```

---

## ii) Compute the average stock price for each terminal node of the tree 

```{r}
# Using forward induction to have cumulative sum C_t along each path, chapter 2 slide 28
#2^25 node to compute = way too much for R -> 

states <- data.frame(
  S      = S0,
  C      = S0,
  weight = 1.0      # physical probability weight (p = 0.5)
)

for (step in 1:n) {
  up <- data.frame(
    S      = states$S * u,
    C      = states$C + states$S * u,
    weight = states$weight * p
  )
  down <- data.frame(
    S      = states$S * d,
    C      = states$C + states$S * d,
    weight = states$weight * (1 - p)
  )
  states <- rbind(up, down)
}

# Compute arithmetic average for each terminal path
states$S_bar <- states$C / (n + 1)

cat("\n=== Task (ii): Arithmetic Average Summary ===\n")
cat("Number of terminal paths:", nrow(states), "\n")
cat("Max average price:  $", round(max(states$S_bar), 4), "\n")
cat("Min average price:  $", round(min(states$S_bar), 4), "\n")
cat("Mean of averages:   $", round(mean(states$S_bar), 4), "\n")

averages_list <- sort(states$S_bar, decreasing = TRUE)


# Plot 2: Distribution of Terminal prices, Using binomial formula with p = 0.5 (physical probability)

j_vals   <- 0:n
prices_T <- S0 * u^j_vals * d^(n - j_vals)
probs_p  <- choose(n, j_vals) * (0.5)^n

terminal_df <- data.frame(
  price = prices_T,
  prob  = probs_p
)

p_dist <- ggplot(terminal_df, aes(x = price, y = prob)) +
  geom_col(fill = "grey40", alpha = 0.85, width = diff(range(prices_T))/30) +
  geom_vline(xintercept = S0,
             color = "red", linetype = "dashed", linewidth = 0.8) +
  annotate("text",
           x = S0 + diff(range(prices_T)) * 0.03,
           y = max(probs_p) * 0.93,
           label = paste0("S0 = $", round(S0, 2)),
           color = "red", size = 3.5, hjust = 0) +
  labs(
    title    = "Distribution of Terminal Stock Prices at t = n",
    subtitle = paste0("n = ", n, " periods   |  Physical Probability p = 0.5"),
    x        = "Terminal Stock Price ($)",
    y        = "Probability"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 10, hjust = 0.5, color = "grey40"),
    axis.title       = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

print(p_dist)

# PLOT 3: DISTRIBUTION OF TERMINAL AVERAGES ─────────────

p_avg <- ggplot(states, aes(x = S_bar, weight = weight)) +
  geom_histogram(bins = 50, fill = "grey40", alpha = 0.85, color = "white") +
  geom_vline(xintercept = S0,
             color = "red", linetype = "dashed", linewidth = 0.8) +
  annotate("text",
           x = S0 + 1, y = Inf, vjust = 2,
           label = paste0("S\u2080 = $", round(S0, 2)),
           color = "red", size = 3.5, hjust = 0) +
  labs(
    title    = "Distribution of Arithmetic Averages at Maturity",
    subtitle = paste0(nrow(states), " paths   |   p = 0.5"),
    x        = "Arithmetic Average Price ($)",
    y        = "Number of Paths"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 10, hjust = 0.5, color = "grey40"),
    axis.title       = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

print(p_avg)
```

The forward induction works by recognizing that the cumulative sum $C_t = \sum_{j=0}^{t} S_j$ satisfies the simple recursion $C_t = C_{t-1} + S_t$, which means we only need to track the pair $(S_t, C_t)$ at each step rather than the full path history, it is the Markov property of the augmented state space. At each time step, every existing state branches into exactly two children (up and down), carrying forward both the new stock price and the updated cumulative sum, with probability weights that multiply by $p = 0.5$ at each branch. After $n = 25$ steps we have exactly $2^{25}$ states, each representing one unique path through the tree, and the arithmetic average is immediately recovered as $\bar{S}_n = \frac{C_n}{(n+1)}$,  no path information is lost.

---

## iii) Compute the different payoffs.

Define the risk-neutral probability and the discount factor using discrete compounding (because of the binomial model)

```{r set-up function parameter}
# Need to compute the risk neutral probability under which there is no arbitrage opportunity (Fundamental Theorem of Asset Pricing)

# define the one-step risk-free growth factor
R_period <- 1 + (r*delta_t)
# define the risk-neutral probability
q <- (R_period-d)/(u - d)
# define the discount factor (for future return)
discount <- 1 / R_period

# print recap of all value needed for the rest of the code
cat("Daily volatility:         ", sigma_daily, "\n")
cat("Annual volatility:        ", sigma, "\n")
cat("Up factor u:              ", u, "\n")
cat("Down factor d:            ", d, "\n")
cat("One-step growth R_period: ", R_period, "\n")
cat("Risk-neutral prob. q:     ", q, "\n")
cat("Discount factor:          ", discount, "\n")

```

```{r compute the payoff}
# we need to compute the payoff at maturity for every path.
states$payoff <- pmax(states$S - states$S_bar, 0)

# print the recap of the payoff ___________________________
cat("\n Floating-Strike Asian Call Payoff\n")
cat("Number of terminal paths: ", nrow(states), "\n")
cat("Minimum payoff:          $", round(min(states$payoff), 4), "\n")
cat("Maximum payoff:          $", round(max(states$payoff), 4), "\n")
cat("Mean payoff:             $", round(mean(states$payoff), 4), "\n")
cat("Median payoff:           $", round(median(states$payoff), 4), "\n")
cat("Share positive payoffs:   ", round(mean(states$payoff > 0) * 100, 2), "%\n")

# Plot the distribution of the payoff at maturity
ggplot(states, aes(x = payoff)) +
  geom_histogram(bins = 50, fill = "grey40", color = "white", alpha = 0.8) +
  geom_vline(xintercept = mean(states$payoff),
             color = "red", linetype = "dashed", linewidth = 0.8) +
  annotate("text",
           x = mean(states$payoff) + 1,
           y = Inf, vjust = 2,
           label = paste0("Mean = $", round(mean(states$payoff), 2)),
           color = "red", size = 3.5, hjust = 0) +
  labs(
    title = "Distribution of Potential Payoffs at Maturity",
    subtitle = "Unweighted across all terminal paths",
    x = "Payoff at maturity",
    y = "Frequency"
  ) +
  theme_minimal()

# Jitter plot in order to showcase how the terminal payoff of the option is related to the average price of the stock of AAPL

set.seed(42)
states_sample <- states[sample(nrow(states), 80000), ]

ggplot(states_sample, aes(x = S, y = payoff)) +
  geom_jitter(alpha = 0.03, size = 0.3, color = "red",
              width = 5, height = 0) +
  labs(
    title    = "Payoff vs Terminal Stock Price",
    subtitle = "Horizontal spread reveals path dependence:\n same Sn, different averages",
    x = expression(Terminal~stock~price~S[n]),
    y = "Payoff ($)"
  ) +
  theme_minimal()
```



## i.v) Weighted Outcome under risk-neutral probability

Use the weighted probability under the no-arbitrage opportunity assumption.

Until now we have determined that a stock would go up or down with the probability of $p =\frac{1}{2}$. However this probability does not excude the possibility of arbitrage where free money can be made without taking any risk. Indeed as the stock move up or down, its return should tend towards 0 on average. However in the same time, a bond could return $0.02\%$. This means that it is possible for an investor to short the stock and invest the amount received now into a bond. Then in the future, he sells the bonds, earn the $0.02\%$ return from the bonds and buy back the share. So in the case where $p = 0.5$, the bond always pays more than what the stock is expected to return which creates an arbitrage opportunity which can be exploited to generate risk-free profit.

Now in order to avoid this situation and guarantee that all assets have the same expected return, one need to adjust the probability. We are thus going to use the probability $q$ under which the stock earns exactly the same amount as the risk-free rate and thus eliminate arbitrage. This is at the core of the Fundamental Assets Pricing Theory presented in Lecture 2 which defines a risk-neutral probability $q$ that serves to guarantee that the stock and the bond combinations have the same expected return that no assets dominates the others, thus resulting in an arbitrage free-market.

In our case, we have an options that pays $X^u$ in the up state and $X^d$ in the down state. The idea is now to find a portfolio compose of a stock $\alpha_1$ and of a bond $beta_1$ that exaclt repliates has the same payoff as our options. This means that in the state up the portfolio $\alpha_1 \cdot S_0 \cdot u + \beta_1 \cdot (1+r) = X^u$ (note that $\alpha_1$ is the part of our initial investment we give to the stock and $\beta_1$ is the part we give to the bonds with $u$ being the effect of the increase of the stock on $S_0$ being the initial value of the stock  and $r$ being the return of the bond). Note that repriprocally, the return in the state down is  $\alpha_1 \cdot S_0 \cdot d + \beta_1 \cdot (1+r) = X^d$. 

Now that both case have been define we need to search for the amount invested in the bond and in the stock $(\alpha_1, \beta_1)$. 

We can rewrite the equation proprely:
$$
\begin{align}
\text{up:}\qquad \alpha_1 \cdot S_0 \cdot u + \beta_1 \cdot (1+r) &= X^u \tag{1} \\
\text{down:}\qquad \alpha_1 \cdot S_0 \cdot d + \beta_1 \cdot (1+r) &= X^d \tag{2}
\end{align}
$$
In order to find $\alpha_1$ we subtract equation 2 from equation 1 and we obtain (with the bonds part cancelling out):
$$
\alpha_1 \cdot S_0 \cdot u - \alpha_1 \cdot S_0 \cdot d = X^u - X^d
$$
It is possible to simplify the equation by grouping the value of the stock $S_0$ in $t=0$.
$$
\alpha_1 \cdot S_0 (u - d) = X^u - X^d
$$
Before simply isolating $\alpha_1$ and finally obtaining:
$$
\boxed{\alpha_1 = \frac{X^u - X^d}{S_0(u - d)}} \tag{3}
$$
The same operation can be conducted for the share invested in the bonds as we now plug equation 3 (our $\alpha_1$) into the initial equation 1.
$$
\begin{aligned}
\frac{X^u - X^d}{S_0(u - d)} \cdot S_0 \cdot d + \beta_1(1+r) \quad = \quad X^d \\ \frac{{X^u - X^d}\cdot S_0 \cdot d}{S_0(u - d)} + \beta_1(1+r)\quad = \quad X^d
\end{aligned}
$$
One can see that the $S_0$ simplifies and that we obtain:
$$
\frac{(X^u - X^d) \cdot d}{u - d} + \beta_1(1+r) = X^d
$$
We then isolate the share invested in $\beta_1$:
$$

\beta_1(1+r) \quad= \quad X^d - \frac{(X^u - X^d) \cdot d}{u - d}\\
\beta_1(1+r) \quad= \quad \frac{X^d(u - d) - (X^u - X^d) \cdot d}{u - d}

$$
In order to simplify we expand the numerator and $X^d \cdot d$ cancels out.
$$
= \frac{X^d \cdot u - X^d \cdot d - X^u \cdot d + X^d \cdot d}{u - d}
= \frac{X^d \cdot u - X^u \cdot d}{u - d}
$$
We can divide it by $(1+r)$ and we obtain $beta_1$.

$$
\boxed{\beta_1 = \frac{1}{1+r} \cdot \frac{X^d \cdot u - X^u \cdot d}{u - d}}
$$


```{r difference between both probability q and p}
# plot the distribution of terminal value under both probability in order to compare them
j_vals   <- 0:n
prices_T <- S0 * u^j_vals * d^(n - j_vals)

# Physical distribution (p = 0.5)
probs_p <- choose(n, j_vals) * p^j_vals * (1 - p)^(n - j_vals)

# Risk-neutral distribution (q)
probs_q <- choose(n, j_vals) * q^j_vals * (1 - q)^(n - j_vals)

dist_df <- data.frame(
  price = rep(prices_T, 2),
  prob  = c(probs_p, probs_q),
  measure = rep(c("Physical (p = 0.5)", "Risk-neutral (q)"), each = n + 1)
)

ggplot(dist_df, aes(x = price, y = prob, fill = measure)) +
  geom_col(position = "dodge", alpha = 0.7, width = diff(range(prices_T)) / 35) +
  geom_vline(xintercept = S0, linetype = "dashed", color = "black", linewidth = 0.5) +
  annotate("text", x = S0 + 5, y = max(c(probs_p, probs_q)) * 0.95,
           label = paste0("S0 = $", round(S0, 2)),
           size = 3.5, hjust = 0) +
  scale_fill_manual(values = c("Physical (p = 0.5)" = "grey40",
                                "Risk-neutral (q)" = "red")) +
  labs(
    title    = "Terminal Price Distribution: Physical vs Risk-Neutral",
    subtitle = paste0("q = ", round(q, 4), " vs p = 0.5 — risk-neutral shifts mass rightward"),
    x = expression(Terminal~stock~price~S[n]),
    y = "Probability",
    fill = "Measure"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```


## v) Estimate the pricing option: (use backward induction)
```{r option value using backward induction}
# we have each terminal payoff
values <- states$payoff 

# in order to use backward induction: we need to start from the payoff at maturity and then loops through each iteration until we reach time 0 (starting point)

for (step in n:1) {
  half   <- length(values) / 2
  # split the tree in two parts (the part above is all the value that ended with an up and the part below is all the down)
  v_up   <- values[1:half]
  v_down <- values[(half + 1):length(values)]
  values <- discount * (q * v_up + (1 - q) * v_down)
  # we compute the probability of having the up and the down before discounting them by the period factor
  # Loop through all the iteration of the tree in order to reach the starting point (unique V0)
}

V0_b<- values
cat(" Option price at t = 0 with backward induction: $", round(V0_b,4))

```

```{r option value using direct forward calculation}
# Control if both model gives the same value:

# find the number of up moves for each path
states$up_moves <- round((log(states$S / S0) - n * log(d)) / (log(u) - log(d))) 

# Risk-neutral probability of each individual path
states$rn_weight <- q^states$up_moves * (1 - q)^(n - states$up_moves)

# Expected payoff under the risk-neutral measure
expected_payoff <- sum(states$rn_weight * states$payoff)

# Discount the expected payoff back to time 0
V0_f<- discount^n * expected_payoff

cat(" The option price at T = 0 with direct forward calculation: $", round(V0_f,4))

```












## vi) Robustness check:

The annualised volatility sigma is the only parameter in our model that we had to estimated from the data rather than simply taking it from the assignment. We compute it as sigma = sigma_daily x sqrt(250), where sigma_daily is the sample standard deviation of the daily log returns over the period 2020-2026. However, this raises a natural question: would our baseline price of 13.89 change significantly if we had chosen a different historical window to estimate sigma?

To answer this, we re-estimate sigma over four alternative sub-periods and re-price the option each time, keeping all other parameters fixed at their baseline values. The four windows are the full 2020-2026 sample, a post-COVID window starting in 2021, a recent three-year window from 2023 to 2026, and the most recent twelve months. The results show that the baseline price is sensitive to whether we include the COVID period of 2020 or not. Excluding it brings sigma down from 0.313 to as low as 0.253, and the option price falls from 13.89 to 11.30. This makes intuitive sense: a lower volatility means a narrower binomial tree, which reduces the expected payoff E_Q[max(Sn - Sn_bar, 0)] and therefore the price. On the other hand, when we use only the most recent twelve months, sigma recovers to 0.309 and the price returns to 13.70, close to our baseline. This is consistent with the elevated market uncertainty observed in early 2026.

We therefore conclude that our baseline price is robust to recent market conditions but sensitive to the inclusion of the COVID volatility spike of 2020. The key modelling assumption being tested here is whether the 2020-2026 window is representative of the volatility that will prevail over the next six months, and there is no objectively correct answer to that question.

```{r robustness check}

raw <- data
raw$AAPL <- raw$Price

# Helper: estimate σ from a data sub-window and re-price
# Uses the same forward enumeration + risk-neutral pricing as Tasks ii/iii/v
reprice_window <- function(data_subset, S0, r, T, n, p = 0.5) {
  log_ret   <- diff(log(data_subset$AAPL))
  sig_daily <- sd(log_ret, na.rm = TRUE)
  sig_ann   <- sig_daily * sqrt(250)

  # Recompute tree parameters with the new sigma
  dt    <- T / n
  u_new <- exp( sig_ann * sqrt(dt))
  d_new <- exp(-sig_ann * sqrt(dt))
  R_new <- exp(r * dt)
  q_new <- (R_new - d_new) / (u_new - d_new)

  # Forward enumeration of all 2^n paths (same logic as Task ii)
  states <- data.frame(S = S0, C = S0, weight = 1.0)
  for (step in 1:n) {
    up   <- data.frame(S      = states$S * u_new,
                       C      = states$C + states$S * u_new,
                       weight = states$weight * p)
    down <- data.frame(S      = states$S * d_new,
                       C      = states$C + states$S * d_new,
                       weight = states$weight * (1 - p))
    states <- rbind(up, down)
  }

  # Payoff and risk-neutral pricing (same logic as Task v)
  states$S_bar    <- states$C / (n + 1)
  states$payoff   <- pmax(states$S - states$S_bar, 0)
  states$up_moves <- round((log(states$S / S0) - n * log(d_new)) /
                             (log(u_new) - log(d_new)))
  states$rn_weight <- q_new^states$up_moves * (1 - q_new)^(n - states$up_moves)

  price <- exp(-r * T) * sum(states$rn_weight * states$payoff)
  list(sigma_daily = sig_daily, sigma_ann = sig_ann, price = price)
}

# Define the four estimation windows
windows <- list(
  list(start = "2020-01-01", end = "2026-12-31", label = "Full window (2020-2026)"),
  list(start = "2021-01-01", end = "2026-12-31", label = "Post-COVID (2021-2026)"),
  list(start = "2023-01-01", end = "2026-12-31", label = "Recent 3 years (2023-2026)"),
  list(start = "2025-04-01", end = "2026-12-31", label = "Recent 12 months")
)

# Run for each window
window_results <- data.frame(
  Window       = character(),
  N_obs        = integer(),
  sigma_daily  = numeric(),
  sigma_ann    = numeric(),
  option_price = numeric(),
  stringsAsFactors = FALSE
)

for (w in windows) {
  sub <- subset(raw, Date >= as.Date(w$start) & Date <= as.Date(w$end))
  sub <- sub[order(sub$Date), ]
  res <- reprice_window(sub, S0, r, T, n)
  window_results <- rbind(window_results, data.frame(
    Window       = w$label,
    N_obs        = nrow(sub) - 1,
    sigma_daily  = res$sigma_daily,
    sigma_ann    = res$sigma_ann,
    option_price = res$price,
    stringsAsFactors = FALSE
  ))
}

# Print results
print(window_results)

cat(sprintf("\nσ range:     [%.4f, %.4f]  (Δ = %.4f)\n",
            min(window_results$sigma_ann),
            max(window_results$sigma_ann),
            diff(range(window_results$sigma_ann))))
cat(sprintf("Price range: [%.4f, %.4f]  (Δ = %.4f USD)\n",
            min(window_results$option_price),
            max(window_results$option_price),
            diff(range(window_results$option_price))))

# Plot
ggplot(window_results, aes(x = option_price, y = Window)) +
  geom_segment(aes(x    = min(option_price) * 0.99,
                   xend = option_price,
                   y    = Window, yend = Window),
               colour = "grey70", linewidth = 0.6) +
  geom_point(aes(colour = sigma_ann), size = 5) +
  geom_text(aes(label = sprintf("$%.2f  (sigma = %.3f)",
                                option_price, sigma_ann)),
            hjust = -0.15, vjust = -0.8, size = 3.2) +
  scale_colour_gradient(low = "#2166ac", high = "#d6604d",
                        name = "sigma (annual)",
                        breaks = c(0.26, 0.28, 0.30, 0.313),
                        labels = c("0.260", "0.280", "0.300", "0.313")) +
  scale_x_continuous(
    labels = scales::dollar_format(),
    expand = expansion(mult = c(0.02, 0.45))
  ) +
  labs(
    title    = "Option Price Sensitivity to sigma Estimation Window",
    subtitle = paste0("Each point is V0 re-priced using sigma from a different sub-period.\n",
                      "r = 1%, S0 = $", round(S0, 2), ", T = 0.5, n = 25."),
    x = "Option price at t = 0 (USD)",
    y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title    = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, colour = "grey40"),
    legend.position = "bottom",
    legend.key.width = unit(2, "cm")
  )
```

