---
title: "Stochastic Method in Finance"
subtitle: "Assignement: Exotic Options"
author: "Antoine Marie - 24-613-762"
format: 
  html: 
    embed-resources: true  
    self-contained: true
    self-contained-math: true
    math-method: mathjax
    
highlight-style: dracula
editor: 
  markdown: 
    wrap: 72
---
```{r setup, echo = FALSE, eval = TRUE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, eval = TRUE)
```

# R Setup
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
```

# 0) Data Preparation

##── 1. Download Apple Data 

```{r}
getwd()
setwd("C:/Users/willi/OneDrive/Uni St.Gallen/Lectures/2nd Year Bachelor/2nd Semester/Stochasitc Method in Finance/Assignment")
data <- read_excel("AAPL.xlsx")
```
Added to clean the environment:
rm(list = ls())

# ── 2. Cleaning 

```{r}
# Convert date to Date format
data$Date <- as.Date(data$Date)

# Rename AAPL.US to Price
colnames(data)[colnames(data) == "AAPL.US"] <- "Price"

# Convert Price to numeric (in case it was read as character)
data$Price <- as.numeric(data$Price)

# Sort ascending (oldest first)
data <- data[order(data$Date), ]

# Filter to 2020 onwards
data <- data[data$Date >= as.Date("2020-01-01"), ]

# Drop unnecessary columns (keep only Date and Price)
data <- data[, c("Date", "Price")]

# Remove any rows with NA
data <- na.omit(data)
```
# ── 3. Verification 
```{r}
cat("Date range:  ", format(min(data$Date)), "to", format(max(data$Date)), "\n")
cat("Observations:", nrow(data), "\n")
cat("Date class:  ", class(data$Date), "\n")
cat("Price class: ", class(data$Price), "\n")
cat("Any NAs:     ", anyNA(data), "\n")

head(data)
tail(data)
```

---

## i) Binomial Tree

# ── Given Parameters 
```{r}
n <- 25           # number of periods
r <- 0.01         # interest rate per annum
S0 <- 267.6100    # Close value on the 27th April 2026 
p <- 0.5          # probability of a up or down
T <- 6/12         # anualise the maturity
delta_t <- T/n    # go from discrete to continuous Random walk to Brownian motion

```

# ── Computations of parameters 
```{r}
#Daily log returns 
returns <- diff(log(data$Price)) #Calculate the daily log returns, slide 

sigma_daily <- sd(returns, na.rm = TRUE) #Calculate the standard deviation, remove na values

sigma <- sigma_daily * sqrt(250) #calculate the annualized volatility, given in assignment

#Up and down factors 
u <- exp(sigma * sqrt(delta_t))  #page 2
d <- exp(-sigma * sqrt(delta_t))
q <- (1+r-d)/(u-d)               # why do we use discret instead of continous formula for the q
```

# ── The binomial tree (matrix format) 
```{r}
# Initialize price tree
stock_tree <- matrix(0, nrow = n + 1, ncol = n + 1)

for (i in 0:n) {        # i = time step
  for (j in 0:i) {      # j = number of up moves
    stock_tree[i - j + 1, i + 1] <- S0 * u^j * d^(i - j)
  }
}

print(stock_tree)
#middle node at maturity [13;25] = S0

# All non-zero terminal nodes
terminal <- stock_tree[stock_tree[, n + 1] != 0, n + 1]

cat("# of terminal nodes:", length(terminal), "\n")
cat("Maximum price:           ", max(terminal), "\n")
cat("Minimum price:           ", min(terminal), "\n")
cat("Middle node (~S0):       ", median(terminal), "\n")

# All terminal prices
cat("\nAll terminal prices:\n")
print(round(terminal, 2))

# Binonmial tree visual 
tree_nodes <- data.frame()
for (i in 0:4) {
  for (j in 0:i) {
    price       <- S0 * u^j * d^(i - j)
    is_reconcile <- (i == 2 & j == 1)
    tree_nodes  <- rbind(tree_nodes, data.frame(
      t         = i,
      y         = j - i/2,
      price     = round(price, 2),
      reconcile = is_reconcile
    ))
  }
}

tree_edges <- data.frame()
for (i in 0:3) {
  for (j in 0:i) {
    y_current <- j - i/2
    tree_edges <- rbind(tree_edges, data.frame(
      x = i, xend = i + 1,
      y = y_current, yend = (j + 1) - (i + 1)/2
    ))
    tree_edges <- rbind(tree_edges, data.frame(
      x = i, xend = i + 1,
      y = y_current, yend = j - (i + 1)/2
    ))
  }
}

p_tree <- ggplot() +
  geom_segment(data = tree_edges,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "grey40", linewidth = 0.6) +
  geom_point(data = tree_nodes[!tree_nodes$reconcile, ],
             aes(x = t, y = y),
             size = 16, color = "grey40") +
  geom_point(data = tree_nodes[tree_nodes$reconcile, ],
             aes(x = t, y = y),
             size = 16, color = "red") +
  geom_text(data = tree_nodes,
            aes(x = t, y = y, label = paste0("$", price)),
            color = "white", size = 2.8, fontface = "bold") +
  annotate("text",
           x = tree_nodes$t[tree_nodes$reconcile] + 0.12,
           y = tree_nodes$y[tree_nodes$reconcile] - 0.42,
           label = "S0 * u * d = S0",
           color = "red", size = 3.2, hjust = 0) +
  annotate("text",
           x = 0:4, y = rep(-2.9, 5),
           label = paste0("t = ", 0:4),
           size = 3.2, color = "grey40") +
  scale_x_continuous(limits = c(-0.3, 5.0)) +
  scale_y_continuous(limits = c(-3.2, 2.6)) +
  labs(
    title    = "Binomial Tree — First 4 Periods",
    subtitle = paste0("S0 = $", round(S0, 2),
                      "   u = ", round(u, 4),
                      "   d = ", round(d, 4),
                      "   p = ", p),
    caption  = "Red node: recombination S0 * u * d = S0"
  ) +
  theme_void() +
  theme(
    plot.title    = element_text(size = 14, face = "bold",
                                 hjust = 0.5, margin = margin(b = 5)),
    plot.subtitle = element_text(size = 10, hjust = 0.5,
                                 color = "grey40", margin = margin(b = 15)),
    plot.caption  = element_text(size = 9, hjust = 0.5,
                                 color = "red", margin = margin(t = 10)),
    plot.margin   = margin(15, 15, 15, 15)
  )

print(p_tree)
```

---

## ii) Compute the average stock price for each terminal node of the tree 

```{r}
# Using forward induction to have cumulative sum C_t along each path, chapter 2 slide 28
#2^25 node to compute = way too much for R -> 

states <- data.frame(
  S      = S0,
  C      = S0,
  weight = 1.0      # physical probability weight (p = 0.5)
)

for (step in 1:n) {
  up <- data.frame(
    S      = states$S * u,
    C      = states$C + states$S * u,
    weight = states$weight * p
  )
  down <- data.frame(
    S      = states$S * d,
    C      = states$C + states$S * d,
    weight = states$weight * (1 - p)
  )
  states <- rbind(up, down)
}

# Compute arithmetic average for each terminal path
states$S_bar <- states$C / (n + 1)

cat("\n=== Task (ii): Arithmetic Average Summary ===\n")
cat("Number of terminal paths:", nrow(states), "\n")
cat("Max average price:  $", round(max(states$S_bar), 4), "\n")
cat("Min average price:  $", round(min(states$S_bar), 4), "\n")
cat("Mean of averages:   $", round(mean(states$S_bar), 4), "\n")

averages_list <- sort(states$S_bar, decreasing = TRUE)


# Plot 2: Distribution of Terminal prices, Using binomial formula with p = 0.5 (physical probability)

j_vals   <- 0:n
prices_T <- S0 * u^j_vals * d^(n - j_vals)
probs_p  <- choose(n, j_vals) * (0.5)^n

terminal_df <- data.frame(
  price = prices_T,
  prob  = probs_p
)

p_dist <- ggplot(terminal_df, aes(x = price, y = prob)) +
  geom_col(fill = "grey40", alpha = 0.85, width = diff(range(prices_T))/30) +
  geom_vline(xintercept = S0,
             color = "red", linetype = "dashed", linewidth = 0.8) +
  annotate("text",
           x = S0 + diff(range(prices_T)) * 0.03,
           y = max(probs_p) * 0.93,
           label = paste0("S0 = $", round(S0, 2)),
           color = "red", size = 3.5, hjust = 0) +
  labs(
    title    = "Distribution of Terminal Stock Prices at t = n",
    subtitle = paste0("n = ", n, " periods   |  Physical Probability p = 0.5"),
    x        = "Terminal Stock Price ($)",
    y        = "Probability"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 10, hjust = 0.5, color = "grey40"),
    axis.title       = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

print(p_dist)

# PLOT 3: DISTRIBUTION OF TERMINAL AVERAGES ─────────────

p_avg <- ggplot(states, aes(x = S_bar, weight = weight)) +
  geom_histogram(bins = 50, fill = "grey40", alpha = 0.85, color = "white") +
  geom_vline(xintercept = S0,
             color = "red", linetype = "dashed", linewidth = 0.8) +
  annotate("text",
           x = S0 + 1, y = Inf, vjust = 2,
           label = paste0("S\u2080 = $", round(S0, 2)),
           color = "red", size = 3.5, hjust = 0) +
  labs(
    title    = "Distribution of Arithmetic Averages at Maturity",
    subtitle = paste0(nrow(states), " paths   |   p = 0.5"),
    x        = "Arithmetic Average Price ($)",
    y        = "Number of Paths"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 10, hjust = 0.5, color = "grey40"),
    axis.title       = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

print(p_avg)
```

The forward induction works by recognizing that the cumulative sum $C_t = \sum_{j=0}^{t} S_j$ satisfies the simple recursion $C_t = C_{t-1} + S_t$, which means we only need to track the pair $(S_t, C_t)$ at each step rather than the full path history, it is the Markov property of the augmented state space. At each time step, every existing state branches into exactly two children (up and down), carrying forward both the new stock price and the updated cumulative sum, with probability weights that multiply by $p = 0.5$ at each branch. After $n = 25$ steps we have exactly $2^{25}$ states, each representing one unique path through the tree, and the arithmetic average is immediately recovered as $\bar{S}_n = \frac{C_n}{(n+1)}$,  no path information is lost.

---

## iii) Compute the different payoffs.

Redefine the probability using the continous function instead of the discrete variable.

```{r set-up function parameter continous}
# define the one-step risk-free growth factor
R <- exp(r * delta_t)
# define the risk-neutral probability
q <- (R-d)/(u - d)
# define the discount factor 
discount <- exp(-r * delta_t)


# print recap of all value needed for the rest of the code
cat("Daily volatility:      ", sigma_daily, "\n")
cat("Annual volatility:     ", sigma, "\n")
cat("Up factor u:           ", u, "\n")
cat("Down factor d:         ", d, "\n")
cat("One-step growth R:     ", R, "\n")
cat("Risk-neutral prob. q:  ", q, "\n")
cat("Discount factor:       ", discount, "\n")
```

```{r compute the payoff}
# we need to compute the payoff at maturity for every path.
states$payoff <- pmax(states$S - states$S_bar, 0)

# print the recap of the payoff ___________________________
cat("\n=== Task (iii): Floating-Strike Asian Call Payoff ===\n")
cat("Number of terminal paths: ", nrow(states), "\n")
cat("Minimum payoff:          $", round(min(states$payoff), 4), "\n")
cat("Maximum payoff:          $", round(max(states$payoff), 4), "\n")
cat("Mean payoff:             $", round(mean(states$payoff), 4), "\n")
cat("Share positive payoffs:   ", round(mean(states$payoff > 0) * 100, 2), "%\n")

# Plot the distribution of the payoff at maturity _________
ggplot(states, aes(x = payoff)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white", alpha = 0.8) +
  labs(
    title = "Distribution of Potential Payoffs at Maturity",
    subtitle = "Unweighted across all terminal paths",
    x = "Payoff at maturity",
    y = "Frequency"
  ) +
  theme_minimal()



head(states)
```

## i.v) Weighted Outcome under risk-neutral probability
Use the weighted probability under the no-arbitrage opportunity. The risk-neutral probability is chosen in order for the expected one-step stock price grows at the risk-free rate (no-arbitrage).


## v) Estimate the pricing option:
We have define all the terminal nods of binamial three. However we build it using the probability that ups and downs are both equal to 1/2. In order to be able to price the options we need to use the risk-neutral path probabilities using q. As a reminder, q is the risk-neutral probability (under the assumption that there is no arbitrage) and we know that all the outcome of the three can be represented as the terminal stock price at maturity multiplicated by the probability of up $u^j$ and the probability of down $d^{n-j}$, noting that $j$ is the number of up in the path of the value of the stock and $n$ is the total number of period which in our case is 25.

```{r option value}
# find the number of up moves for each path
states$up_moves <- round((log(states$S / S0) - n * log(d)) / (log(u) - log(d))) 

# Risk-neutral probability of each individual path
states$rn_weight <- q^states$up_moves * (1 - q)^(n - states$up_moves)

# Expected payoff under the risk-neutral measure
expected_payoff <- sum(states$rn_weight * states$payoff)

# Discount the expected payoff back to time 0
V0 <- exp(-r * T) * expected_payoff

print(V0)

# Payoff plot
states |>
  group_by(up_moves) |>
  summarise(
    terminal_price = mean(S),
    average_payoff = mean(payoff),
    .groups = "drop"
  ) |>
  ggplot(aes(x = terminal_price, y = average_payoff)) +
  geom_col(fill = "grey40", alpha = 0.85) +
  geom_vline(xintercept = S0, color = "red", linetype = "dashed") +
  labs(
    title = "Average Payoff by Terminal Stock Price",
    subtitle = "Floating-strike Asian call option",
    x = "Terminal stock price",
    y = "Average payoff"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal()
```



